How do changes in ex vary between countries?

### 2022-03-07 draft_paper.Rmd
changed to null_model
get optim to work; realise I probably shouldn't have tried!

# Here 2015 has the lowest value, but this might be an artefact of the data period covered. 
# We can also see that 2012 is the next best candidate.
```

>>>>

# Here 2015 has the lowest value, but this might be an artefact of the data period covered. 
# We can also see that 2012 is the next best candidate.

## Generalising and automating.

# We've manually found the best break points, and compared the best alt with the standard null model 
# for each of these. What we should do now is standardise the process so it 'works' automatically 
# for each of the population subgroups.

# As with creating synthetic germany, the grid/line search approach shown above should work, 
# but I'm going to try to use optim nonetheless for pedagogic and perfectionist(/masochistic) reasons

# optim uses some starting conditions (and sometimes other conditions) to use an algorithm 
# to find an input or group of inputs that give the best possible output from a function that takes the 
# input/inputs.

# best can either mean the lowest or the highest achievable value. When trying to minimise errors
# (as with model specifications) that usually means trying to get the minimum value.

# So, the input to the function will be the tau value. 
# The tau value range should have some reasonable limits. These can be specified 
# for some kinds of algorithm optim uses.
# the limits should also be determined by the data. I don't think the tau should be 
# set at the first or last observed year, as it doesn't make sense to speak of a 'trend' 
# from a single observation. This thinking generalises to at least two observations too. 
# So, that implies there should be a 'buffer' within which the optim function will look 
# for possible tau values. This should be an adjustable parameter for passing for 
# optim, but not the parameter for optim to determine.

# There's a special type of optim algorithm for bounded optimisation like this.
# 'L-BFGS-B'. This can be selected using the method argument, and also 
# requires the lower and upper arguments be specified, for the lower 
# and upper values to search within respectively.

# so, we need to write a function a bit like the following

get_best_tau_from_optim <- function(df, buffer = 2, what = c("tau", "optim")){
  # We can put in some data validation checks too

  what <- match.arg(what)

  years <- df$year 
  tau_lower <- min(years) + buffer # the lower bound 
  tau_upper <- max(years) - buffer # the upper bound
  tau_start <- (tau_upper + tau_lower) / 2 # start just in the middle 
  # (This shouldn't matter if the algorithm is robust)

  get_result_from_model <- function(par, df, what = c("AIC", "BIC", "model")){
    # Again, can carry out more validation checks 
    what <- match.arg(what)

    tau <- par[["tau"]] # This is how parameters are packed up by optim

    model <-
      df %>% 
        arrange(year) %>% 
        mutate(delta_ex = ex - lag(ex)) %>% 
        filter(!is.na(delta_ex)) %>% 
        mutate(T_param = ifelse(year < tau, FALSE, TRUE)) %>% 
        lm(delta_ex ~ T_param, data = .)

    if (what == "model") {return(model)      }
    if (what == "BIC")   {return(BIC(model)) }
    if (what == "AIC")   {return(AIC(model)) }

    # NULL should never be returned! If it has something's gone wrong!
    NULL
  }

  optim_obj <- 
    optim(
      par = list(tau = tau_start), 
      fn = get_result_from_model, 
      method = "L-BFGS-B",
      lower = tau_lower,
      upper = tau_upper,
      df = df
    )

  # Need to make the result dependent on the what argument

  if (what == "tau") {return(optim_obj[["par"]])}
  else if (what == "optim"){ return(optim_obj) }

  # Again, the following should never be triggered
  NULL
}

# Let's see if this now works in the functional programming context, using 
# nest and tibble elements

hmd_ex_selected_countries_with_synth %>%
  filter(year >= 1979) %>% 
  group_by(code, x, sex) %>% 
  nest() %>% 
  mutate(
    best_tau = map_dbl(data, get_best_tau_from_optim)
  )

# I think this is getting stuck on local optima, so instead maybe 
# the more exhaustive search should be used after all!
